EoS-insensitive relations after GW170817

### 2019-02-06 paper/bin-love-after-gw170817.tex
added responses to comments.

{\ny{This is a bit strange, but probably still correct. But ``this'' I mean the fact that when you fit only to the hybrid stars you do not get better than $1\%$ universality. If I look at the hybrid curves alone, they look like the same simple lines as in the neutron star case (but of course a $1\%$ difference wouldn't show on this scale). Why is it that we can lower fractional difference in the NS case but not in the hybrid case?}}
{\zc{Nico, I think the reason for this is because there is nothing ``constrained" about the hybrid stars - these are basically ``unconstrained" EoSs so we can't really reduce the fractional difference any more.}}
{\ny{Can I see the relative fractional different between each fit and the data for hybrid stars only? (ie. the analogous thing to the green dashed lines of the figure below, but also with the fits to the total set, to just the hybrids and to just the neutron stars.) I wonder whether there is any improvement at all when you fit just to hybrid stars or whether there are any features.}}
{\zc{Nico, Yes of course! I've added a new figure (for example only - this can be removed later) - Fig.~\ref{fig:example} showing the fractional difference for hybrid stars only for the three cases shown in the text. Looking at the figures, they don't change much at all in shape for the first two cases - just a minor vertical shift. The third case (fits to hybrids only) looks different but not by much.}}
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{compareHybrid.pdf}
\end{center}
\caption{
\zc{Example figure applying to Nicos comment on the differences between the three different cases of Hybrid star fits: Case I - Complete set of 100 constrained EoSs combined
with the 10 hybrid star EoSs, Case II - Complete set of 100 constrained EoSs alone, Case III - Complete set of 10 hybrid star EoSs alone.}
}
\label{fig:example}
\end{figure}

@@

\kc{The R-Love fit is not a new relation, it is the same as C-Love. So instead of refitting you should just reexpress the C(Love) fit as R=m/C(Love). The errors you are finding here are way larger than error I've been finding from simply using C(Love) to compute the radius, and this might be the reason.}
In addition, Fig.~\ref{fig:rlove} displays the relationship between NS radius and tidal deformability for the constrained and unconstrained sets of EoSs.
Similar to the fits performed for the I-Love-Q and C-Love relations, we fit the R-Love relations to Eq.~\ref{eq:ILQfitNew}.
In this case, we utilize the same Newtonian control factor found in Eq.~\ref{eq:cloveFit} for the C-Love relations, substituting the NS mass with the polytropic form $M\propto R^{(3-n)/(1-n)}$~\cite{Shapiro:MRpolytrope}.
Noting that most physical EoSs obey approximate polytropic indices between $0.50$ and $1.00$, as shown in Ref.~\cite{Yagi:binLove}, we choose a central value of $n=0.75$ for use in this analysis, resulting in $M\propto R^9$.
Absorbing the proportionality constant into $K_{R\Lambda}$ gives the approximate Newtonian relationship as:
\begin{equation}
R^N \approx K_{R\Lambda}\Lambda^{-1/40}.
\end{equation}
The results of this fit are displayed in Tab.~\ref{tab:ILQfitNew}, and the maximal EoS variability is again shown in Tab.~\ref{tab:maxVar}.
Previously, this relation observed large relative uncertainties on the order of $\sim50\%$ on the mapping from tidal deformability to radius, as can be seen by the unconstrained EoSs in Fig.~\ref{fig:rlove}.
However, upon the restriction to soft EoSs conforming to GW170817, we observe a large suppression in EoS variability to the order of $\sim20\%$.
This results in new EoS-insensitive relations, with relative uncertainties on the same scale as found in the binary Love relations discussed below in Sec.~\ref{sec:binary}.
\begin{figure}
\begin{center} 
\includegraphics[width=1\columnwidth]{RL.eps}
\end{center}
\caption{
Similar to Fig.~\ref{fig:clove} but for the R-Love relationship between the NS radius and tidal deformability.
In this figure, we also construct fits to  Eq.~\ref{eq:ILQfitNew}, limiting to the Newtonian case $R \sim M\Lambda^{1/5}$, with $M\sim R^{(3-n)/(1-n)}$ for polytropic index $n$.
Observe how the fractional difference from the fit is reduced significantly between the unconstrained and constrained sets of EoSs, shown by the comparisons in Tab.~\ref{tab:maxVar}.
Also shown in this figure is the hybrid star R-Love relations along with their fractional difference from the constrained EoS fits (dashed green). \kc{The radius needs to be in km in both panels (not fractional error).}
}
\label{fig:rlove}
\end{figure}

@@

{\ny{I must have missed something. Why do the unconstrained magenta lines extend way passed the constrained green lines? I didn't see this discussed anywhere.}}
{\zc{This is because I've cut all the EoSs off at masses of 1 M$_{\odot}$, and typically the unconstrained EoSs (maroon) have larger ranges of $\Lambda$ than the unconstrained ones. So when we take $\Lambda_1+\Lambda_2$, we end up getting higher values of $\Lambda_s$ for the unconstrained EoSs than we do for the constrained ones.}}
\begin{figure}
\begin{center} 
\includegraphics[width=\linewidth]{binLoveAbsolute.eps}%{binLove.eps}
\end{center}
\caption{
Binary Love relations shown for the constrained EoSs (dotted maroon) and unconstrained EoSs (solid green) for various values of mass ratio: $q=0.9$, $q=0.75$, and $q=0.50$.
In this figure, the top panel displays the EoS-insensitive relations with fits given by Eq.~\ref{eq:binLovefit} shown by dashed black lines (note there are 2 fits for each mass ratio, corresponding to the constrained and unconstrained sets of EoSs), while the bottom 3 panels correspond to the absolute EoS variation for each mass ratio.
Observe how the constrained set of EoSs show a reduction in EoS variation for both the unconstrained set, and from the results shown in Ref.~\cite{Yagi:binLove}.
The maximal EoS variability (shown as a the fractional difference for comparison purposes) for each case is tabulated and compared in Tab.~\ref{tab:maxVar}.
Additionally shown is the binary Love relations for the 10 hybrid star EoSs (dashed bright green curves), as well as the HS fits (dot-dashed blue curves) discussed in Sec.~\ref{sec:binLove-hybrid}.
Observe the large deviations from the fit as the mass ratio $q$ increases, returning near the fit before the transitional pressure at large values of $\Lambda_s$.
This is shown in the bottom three panels as the absolute difference from the NS fit for comparison, while the HS fits show similar EoS variability to the constrained EoSs.
}
\label{fig:binLove}
\end{figure}

@@

To determine how each of class of binary NS systems obey our improved binary Love relations, we pick a representative set of 10 binary pairs from each class from the ACB5 hybrid EoS, with a transition at $M_{\text{tr}} \approx 1.4 \text{ M}_{\odot}$.
Binary Love values are computed for each pair, and fractional residuals from the binary Love curve found in Sec.~\ref{sec:binLove-nuclear} are determined and compared to the values from the constrained set of EoSs.
The results are shown in Fig.~\ref{fig:hybrid}.
First, we note that the NS/NS combination of stars gives consistent results to that found earlier -- as expected.
Next, we observe that HS/HS combinations agree moderately with previous relations\footnote{Note that these EoSs do not fit within GW170817s 90\% posterior on EoS - so we expect deviations consistent with the ``unconstrained" set of EoSs, rather than the ``constrained" set.}, while HS/NS combinations disagree significantly.
This can be explained by the large reduction in tidal deformability proceeding the transition pressure for hybrid stars, as shown in Fig.~\ref{fig:hybrid}.
When one star lies on the hadronic-matter branch while the other is on the quark-matter branch, 
the difference between tidal deformabilities becomes larger than expected for pure hadronic-matter stars.
This results in a large deviation in the NS binary Love relations $\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$, disrupting the overall universality.
{\ny{Sorry, I don't get it. Why does seeing ``a large difference in Lambda tilde'' explain the loss of universality? We have cases with small enough q that show much more universality than what we find for the same value of q in the HS/NS case. I guess what I am saying is that we should explore this more from a theory side to try to provide an explanation, or we should say nothing (or say we don't know why this is the case).}}
{\zc{Nico, sorry this was a bad explanation on my part. when the two stars each see a large jump in $\Lambda$, the differences and sums of tidal deformabilities also see a big jump, and they ``jump off" the universal line as can be seen by the dashed bright green lines in fig.~\ref{fig:binLove}. I've rewritten this sentence to (hopefully) make more sense.}}

@@

This discrepancy between the hybrid and hadronic matter EoSs can be seen in Fig.~\ref{fig:binLove}, where the binary Love relations for the 10 hybrid star EoSs are shown alongside the constrained and unconstrained nuclear matter EoSs.
Observe how the residuals from the fit are quite good at low pressures (large $\Lambda_s$) where the the binaries consist of NS/NS combinations, before reaching the transitional pressure where the deviations become increasingly large ($\sim160$\% for $q=0.90$, $\sim37$\% for $q=0.75$, and $\sim6$\% for $q=0.50$) for HS/HS and HS/NS binaries. {\ny{Why is there a hump for the HS cases in green in Fig. 7?}}{\zc{The hump comes from the transition between NS and HS branches, as I describe in my above comment. If you are referring to the bottom panels of Fig 7, this hump is because I display the fractional difference of the \emph{HS} EoSs from the \emph{NS} fits - to show how badly they fit (although that is not as visible now that we are plotting absolute difference. Instead, I could plot their fractional difference from their own "HS" fits too. We can discuss this.}}

@@

We conclude this section with the observation that the hybrid star EoSs do \emph{not} conform to the fits performed on nuclear matter EoSs.
This is not surprising, as the hybrid branch of the EoSs forces inconsistent results between $\Lambda_1$ and $\Lambda_2$ with different masses; an issue not present for the single-star I-Love-Q relations.
By separating the two branches of the hybrid EoSs, we find a new ``hybrid" branch binary Love fit which gives consistent results to that found for the nuclear matter fits. 
{\ny{But wait, the Newtonian limit is not valid for this hybrid branch, since you are not allowed to go to low enough central densities, right? If so, what are you using for the controlling factor? I found the description in the 2 paragraphs above a bit muddled and short. It is not exactly clear to me what the message you were trying to send was.}}
{\zc{You're right - I tried to change this by using leaving the polytropic index (n) as an additional fitting parameter. Doing this gives n=1.7, and all other parameters $>10^7$. The fit looks good but I'm skeptical about those fit values. Additionally, if I set n=0, the fits diverge. Instead, I wrote a few sentences in the paper saying how it doesn't properly limit to the Newtonian case, and how we tried to remedy it. I'm welcome to any suggestions!}}

@@

{\ny{Super confused about Fig. 1...}}
{\zc{Yes, there is a \emph{lot} of info in this figure. The filled blue circles are the statistical uncertainties on $\lambda_0$ for GW170817 as if it had been detected on Detector A=(O2, aLIGO, A+, Voyager, ET, CE), where the x-axis corresponds to the SNR on that detector as if it found GW170817. The filled turquoise region is the combined statistical uncertainty on the detection of $N_A$ events on each detector A, computed using the bulleted process above (moved to the appendix). There is some height to this shaded region because I consider the the upper and lower error bars of the local BNS merger rate $\mathcal{R}$ - giving us a range of $N_A$'s to consider for each detector. We can talk more about this over Skype.}}

\kc{me too}

>>>>

{\ny{Can I see the relative fractional different between each fit and the data for hybrid stars only? (ie. the analogous thing to the green dashed lines of the figure below, but also with the fits to the total set, to just the hybrids and to just the neutron stars.) I wonder whether there is any improvement at all when you fit just to hybrid stars or whether there are any features.}}
{\zc{Nico, Yes of course! I've added a new figure (for example only - this can be removed later) - Fig.~\ref{fig:example} showing the fractional difference for hybrid stars only for the three cases shown in the text. Looking at the figures, they don't change much at all in shape for the first two cases - just a minor vertical shift. The third case (fits to hybrids only) looks different but not by much.}}
{\ny{Ok, this makes more sense to me now. The universality does increase, but only by a small amount (which you did say before -- something like from 1.7 to 1 per cent.). Ok, we can remove these figures and these comments if you'd like. }}
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{compareHybrid.pdf}
\end{center}
\caption{
\zc{Example figure applying to Nicos comment on the differences between the three different cases of Hybrid star fits: Case I - Complete set of 100 constrained EoSs combined
with the 10 hybrid star EoSs, Case II - Complete set of 100 constrained EoSs alone, Case III - Complete set of 10 hybrid star EoSs alone.}
}
\label{fig:example}
\end{figure}

@@

\kc{The R-Love fit is not a new relation, it is the same as C-Love. So instead of refitting you should just reexpress the C(Love) fit as R=m/C(Love). The errors you are finding here are way larger than error I've been finding from simply using C(Love) to compute the radius, and this might be the reason.} {\ny{That's a good idea. But it is not immediately obvious that this will help. Kent and I have found that when you look at relations between dimensionless quantities, then the universality is typically stronger than when you use dimensionful quantities (incidentally, this is why the Doneva paper about rapidly-rotating I-Love-Q is wrong, as pointed out by Haris and George).}}
In addition, Fig.~\ref{fig:rlove} displays the relationship between NS radius and tidal deformability for the constrained and unconstrained sets of EoSs.
Similar to the fits performed for the I-Love-Q and C-Love relations, we fit the R-Love relations to Eq.~\ref{eq:ILQfitNew}.
In this case, we utilize the same Newtonian control factor found in Eq.~\ref{eq:cloveFit} for the C-Love relations, substituting the NS mass with the polytropic form $M\propto R^{(3-n)/(1-n)}$~\cite{Shapiro:MRpolytrope}.
Noting that most physical EoSs obey approximate polytropic indices between $0.50$ and $1.00$, as shown in Ref.~\cite{Yagi:binLove}, we choose a central value of $n=0.75$ for use in this analysis, resulting in $M\propto R^9$.
Absorbing the proportionality constant into $K_{R\Lambda}$ gives the approximate Newtonian relationship as:
\begin{equation}
R^N \approx K_{R\Lambda}\Lambda^{-1/40}.
\end{equation}
The results of this fit are displayed in Tab.~\ref{tab:ILQfitNew}, and the maximal EoS variability is again shown in Tab.~\ref{tab:maxVar}.
Previously, this relation observed large relative uncertainties on the order of $\sim50\%$ on the mapping from tidal deformability to radius, as can be seen by the unconstrained EoSs in Fig.~\ref{fig:rlove}.
However, upon the restriction to soft EoSs conforming to GW170817, we observe a large suppression in EoS variability to the order of $\sim20\%$.
This results in new EoS-insensitive relations, with relative uncertainties on the same scale as found in the binary Love relations discussed below in Sec.~\ref{sec:binary}.
\begin{figure}
\begin{center} 
\includegraphics[width=1\columnwidth]{RL.eps}
\end{center}
\caption{
Similar to Fig.~\ref{fig:clove} but for the R-Love relationship between the NS radius and tidal deformability.
In this figure, we also construct fits to  Eq.~\ref{eq:ILQfitNew}, limiting to the Newtonian case $R \sim M\Lambda^{1/5}$, with $M\sim R^{(3-n)/(1-n)}$ for polytropic index $n$.
Observe how the fractional difference from the fit is reduced significantly between the unconstrained and constrained sets of EoSs, shown by the comparisons in Tab.~\ref{tab:maxVar}.
Also shown in this figure is the hybrid star R-Love relations along with their fractional difference from the constrained EoS fits (dashed green). \kc{The radius needs to be in km in both panels (not fractional error).}
}
\label{fig:rlove}
\end{figure}

@@

{\ny{I must have missed something. Why do the unconstrained magenta lines extend way passed the constrained green lines? I didn't see this discussed anywhere.}}
{\zc{This is because I've cut all the EoSs off at masses of 1 M$_{\odot}$, and typically the unconstrained EoSs (maroon) have larger ranges of $\Lambda$ than the unconstrained ones. So when we take $\Lambda_1+\Lambda_2$, we end up getting higher values of $\Lambda_s$ for the unconstrained EoSs than we do for the constrained ones.}}
{\ny{Ahh...yes, makes sense. Please add a sentence or two about this in the text.}}
\begin{figure}
\begin{center} 
\includegraphics[width=\linewidth]{binLoveAbsolute.eps}%{binLove.eps}
\end{center}
\caption{
Binary Love relations shown for the constrained EoSs (dotted maroon) and unconstrained EoSs (solid green) for various values of mass ratio: $q=0.9$, $q=0.75$, and $q=0.50$.
In this figure, the top panel displays the EoS-insensitive relations with fits given by Eq.~\ref{eq:binLovefit} shown by dashed black lines (note there are 2 fits for each mass ratio, corresponding to the constrained and unconstrained sets of EoSs), while the bottom 3 panels correspond to the absolute EoS variation for each mass ratio.
Observe how the constrained set of EoSs show a reduction in EoS variation for both the unconstrained set, and from the results shown in Ref.~\cite{Yagi:binLove}.
The maximal EoS variability (shown as a the fractional difference for comparison purposes) for each case is tabulated and compared in Tab.~\ref{tab:maxVar}.
Additionally shown is the binary Love relations for the 10 hybrid star EoSs (dashed bright green curves), as well as the HS fits (dot-dashed blue curves) discussed in Sec.~\ref{sec:binLove-hybrid}.
Observe the large deviations from the fit as the mass ratio $q$ increases, returning near the fit before the transitional pressure at large values of $\Lambda_s$.
This is shown in the bottom three panels as the absolute difference from the NS fit for comparison, while the HS fits show similar EoS variability to the constrained EoSs.
}
\label{fig:binLove}
\end{figure}

@@

To determine how each of class of binary NS systems obey our improved binary Love relations, we pick a representative set of 10 binary pairs from each class from the ACB5 hybrid EoS, with a transition at $M_{\text{tr}} \approx 1.4 \text{ M}_{\odot}$.
Binary Love values are computed for each pair, and fractional residuals from the binary Love curve found in Sec.~\ref{sec:binLove-nuclear} are determined and compared to the values from the constrained set of EoSs.
The results are shown in Fig.~\ref{fig:hybrid}.
First, we note that the NS/NS combination of stars gives consistent results to that found earlier -- as expected.
Next, we observe that HS/HS combinations agree moderately with previous relations\footnote{Note that these EoSs do not fit within GW170817s 90\% posterior on EoS - so we expect deviations consistent with the ``unconstrained" set of EoSs, rather than the ``constrained" set.}, while HS/NS combinations disagree significantly.
This can be explained by the large reduction in tidal deformability proceeding the transition pressure for hybrid stars, as shown in Fig.~\ref{fig:hybrid}.
When one star lies on the hadronic-matter branch while the other is on the quark-matter branch, 
the difference between tidal deformabilities becomes larger than expected for pure hadronic-matter stars.
This results in a large deviation in the NS binary Love relations $\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$, disrupting the overall universality.

@@

This discrepancy between the hybrid and hadronic matter EoSs can be seen in Fig.~\ref{fig:binLove}, where the binary Love relations for the 10 hybrid star EoSs are shown alongside the constrained and unconstrained nuclear matter EoSs.
Observe how the residuals from the fit are quite good at low pressures (large $\Lambda_s$) where the the binaries consist of NS/NS combinations, before reaching the transitional pressure where the deviations become increasingly large ($\sim160$\% for $q=0.90$, $\sim37$\% for $q=0.75$, and $\sim6$\% for $q=0.50$) for HS/HS and HS/NS binaries. {\ny{Why is there a hump for the HS cases in green in Fig. 7?}}{\zc{The hump comes from the transition between NS and HS branches, as I describe in my above comment. If you are referring to the bottom panels of Fig 7, this hump is because I display the fractional difference of the \emph{HS} EoSs from the \emph{NS} fits - to show how badly they fit (although that is not as visible now that we are plotting absolute difference. Instead, I could plot their fractional difference from their own "HS" fits too. We can discuss this.}} {\ny{I was referring to the hump in the top panel (ie. in the R-Love relation). It'd be good to add a sentence to the text that points this hump out and explains it as you did above.}}

@@

We conclude this section with the observation that the hybrid star EoSs do \emph{not} conform to the fits performed on nuclear matter EoSs.
This is not surprising, as the hybrid branch of the EoSs forces inconsistent results between $\Lambda_1$ and $\Lambda_2$ with different masses; an issue not present for the single-star I-Love-Q relations.
By separating the two branches of the hybrid EoSs, we find a new ``hybrid" branch binary Love fit which gives consistent results to that found for the nuclear matter fits. 
{\ny{But wait, the Newtonian limit is not valid for this hybrid branch, since you are not allowed to go to low enough central densities, right? If so, what are you using for the controlling factor? I found the description in the 2 paragraphs above a bit muddled and short. It is not exactly clear to me what the message you were trying to send was.}}
{\zc{You're right - I tried to change this by using leaving the polytropic index (n) as an additional fitting parameter. Doing this gives n=1.7, and all other parameters $>10^7$. The fit looks good but I'm skeptical about those fit values. Additionally, if I set n=0, the fits diverge. Instead, I wrote a few sentences in the paper saying how it doesn't properly limit to the Newtonian case, and how we tried to remedy it. I'm welcome to any suggestions!}} {\ny{Hmmm....ok.}}

@@

{\ny{Super confused about Fig. 1...}}
\kc{me too}
{\zc{Yes, there is a \emph{lot} of info in this figure. The filled blue circles are the statistical uncertainties on $\lambda_0$ for GW170817 as if it had been detected on Detector A=(O2, aLIGO, A+, Voyager, ET, CE), where the x-axis corresponds to the SNR on that detector as if it found GW170817. The filled turquoise region is the combined statistical uncertainty on the detection of $N_A$ events on each detector A, computed using the bulleted process above (moved to the appendix). There is some height to this shaded region because I consider the the upper and lower error bars of the local BNS merger rate $\mathcal{R}$ - giving us a range of $N_A$'s to consider for each detector. We can talk more about this over Skype.}} 
{\ny{I'm happier with this rewrite (makes more sense to me now).}}
